Inferring Fine-Scale Animal Behaviors using Hidden Markov Models

### 2020-06-05 Report/main.tex
Update simulation study
Many Updates
Updates from Overleaf

\title{Inferring Fine Scale Behaviors Within Hierarchical Hidden Markov Models}
\author{Evan Sidrow}
\date{\today}

@@

In the field of animal movement, recent advances in high-frequency tagging technology have made available a vast amount of data which can exhibit simultaneous behavioral processes occurring at different time scales. One way to model this data is to use a hierarchical hidden Markov model (HHMM), where the system is modeled as a nested structure of hidden Markov models (HMMs). At very short time scales, however, observations can exhibit complicated dependence structures which cannot be easily captured by traditional HMMs. This work investigates how to incorporate fine-scale processes into the larger structure of HHMMs while maintaining computational efficiency. We apply our method to dive data collected from a northern resident killer whale off the coast of British Columbia, Canada.

@@

This work investigates how to incorporate fine-scale processes into the larger structure of hierarchical hidden Markov models while maintaining computational efficiency. We bridge the gap between discrete hidden Markov models and continuous-time stochastic process models by showing that the two are equivalent under certain conditions. In addition, we extract features from highly structured sub-dive behaviors that otherwise could not be modeled with a simple HMM. Finally, we apply our method to dive data collected from a Northern resident killer whale off the coast of British Columbia, Canada.

%\input Sections/intro.tex
\input Sections/HHMM.tex
\input Sections/carHMM.tex
\input Sections/fourier.tex
%\input Sections/simulation.tex
%\input Sections/results.tex
%\input Sections/conclusion.tex

>>>>

\title{Inferring Fine-Scale Animal Behaviors using Hidden Markov Models}
\author{Evan Sidrow}
\date{\today}

@@

@@

This work investigates how to incorporate fine-scale processes into the larger structure of hierarchical hidden Markov models while maintaining computational efficiency. We bridge the gap between discrete hidden Markov models and continuous-time stochastic process models by showing that the two are equivalent under certain conditions. In addition, we extract features from highly structured sub-dive behaviors using signal processing techniques. These features otherwise could not be modeled with a simple HMM. Finally, we apply our method to dive data collected from two Northern resident killer whales off the coast of British Columbia, Canada.

%\input Sections/intro.tex
\input Sections/HHMM.tex
\input Sections/carHMM.tex
\input Sections/fourier.tex
\input Sections/simulation.tex
\input Sections/results.tex
\input Sections/conclusion.tex
